\Large{Adversarial Variational Optimization of Non-Differentiable Simulators}

### 2017-07-13 tex/paper.tex
AVO-EB

\begin{abstract}
Complex computer simulators are increasingly used across fields of science as
generative models tying parameters of an underlying theory to
experimental observations. Inference in this setup is often
difficult, as simulators rarely admit a tractable density or likelihood
function. We introduce Adversarial Variational Optimization (AVO), a likelihood-free
inference algorithm for fitting a non-differentiable generative model to
continuous or discrete  data. \kcnote{empirical variational bayes}. We adapt the training procedure of generative
adversarial networks by replacing the differentiable generative network with a
domain-specific simulator. We solve the resulting non-differentiable
minimax problem by minimizing variational upper bounds of the two adversarial objectives.
Effectively, the procedure results in learning an arbitrarily tight
proposal distribution over simulator parameters, such that the corresponding
marginal distribution of the generated data matches the observations.
We present results of the method with simulators producing both discrete and continuous data.

@@

\subsection{Adversarial variational inference\\
            through model criticism}

@@

Typically in variational inference one starts with a tractable prior
$p(\bftheta)$ and a tractable likelihood $p(\bfx |\bftheta )$, thus fixing the
posterior $p(\bftheta | \bfx )$.   After optimizing $q(\bftheta | \bfpsi)$ to
approximate the posterior, one would criticize the model through the the
posterior predictive distribution as advocated by Box~\citep{box1980sampling}.
Effectively, we are optimizing variational posterior $q(\bftheta | \psi)$ by
directly minimizing the  divergence of the corresponding posterior predictive
distribution $\qxpsi$ and the data distribution $p_r(\bfx)$ without evaluating
the intractable likelihood $p(\bfx | \bftheta )$ or specifying a prior.
Consequently, the proposed algorithm is a form of variational inference through
model criticism~\citep{box1980sampling}, which we denote AVI-MC. While the
likelihood is intractable, it is fixed, thus one can view the optimization of
$\bfpsi$ through the lens of empirical Bayes, though the prior family
$p(\bftheta | \bfpsi)$ is not explicitly specified.

@@

\kcnote{Discuss hierarchical models and iid data x. Factor of $N/M$ for AVI-MC?}

@@

When $\rho$ is the Wasserstein distance, $\bftheta^*$ is referred to as the
minimum Wasserstein estimator. When the model is well specified, $\bftheta^*$
coincides with the true data-generating parameter; however, if the model is
misspecified, $\bftheta^*$ is typically different from the maximum likelihood
estimator (MLE), which is the minimizer of $KL(p_r(\bfx) | p(\bfx |
\bftheta)$~\citep{bernton2017inference}. In order to more effectively target
point estimates $\bftheta^*$, we augment Eqn.~\ref{eqn:vo-ug} with a
regularization term corresponding to the differential entropy $H$ of the
proposal distribution $q(\bftheta | \bfpsi)$. That is,
\begin{equation}
    U_g = \mathbb{E}_{\bftheta \sim q(\bftheta|\bfpsi)} [ {\cal L}_g ] + \gamma H(q(\bftheta|\bfpsi))
\end{equation}
where $\gamma \in \mathbb{R}^+$ is a hyper-parameter controlling the trade-off
between the generator objective and the tightness of the proposal distribution.
For small values of $\gamma$,
proposal distributions with large entropy are not penalized, which results
in learning a smeared variation of the original simulator corresponding to the posterior predictive model. On the other hand,
for large values of $\gamma$, the procedure is encouraged to fit a proposal
distribution with low entropy, which has the effect of concentrating its density
tightly around one or a few $\bftheta$ values.

>>>>

\begin{abstract}
Complex computer simulators are increasingly used across fields of science as
generative models tying parameters of an underlying theory to
experimental observations. Inference in this setup is often
difficult, as simulators rarely admit a tractable density or likelihood
function. We introduce Adversarial Variational Optimization (AVO), a likelihood-free
inference algorithm for fitting a non-differentiable generative model or to perform empirical Bayes through variational inference. 
We adapt the training procedure of generative
adversarial networks by replacing the differentiable generative network with a
domain-specific simulator. We solve the resulting non-differentiable
minimax problem by minimizing variational upper bounds of the two adversarial objectives.
Effectively, the procedure results in learning an arbitrarily tight
proposal distribution over simulator parameters, such that the corresponding
marginal distribution of the generated data matches the observations.
We present results of the method with simulators producing both discrete and continuous data.

@@

\subsection{Empirical Bayes through Adversarial Variational Inference}

@@

Typically in variational inference one starts with a tractable prior
$p(\bftheta)$ and a tractable likelihood $p(\bfx |\bftheta )$, thus fixing the
posterior $p(\bftheta | \bfx )$.   After optimizing $q(\bftheta | \bfpsi)$ to
approximate the posterior, one would criticize the model through the the
posterior predictive distribution as advocated by Box~\citep{box1980sampling}.
Effectively, we are optimizing variational posterior $q(\bftheta | \psi)$ by
directly minimizing the  divergence of the corresponding posterior predictive
distribution $\qxpsi$ and the data distribution $p_r(\bfx)$ without evaluating
the intractable likelihood $p(\bfx | \bftheta )$ or specifying a prior.
Consequently, the proposed algorithm is a form of variational inference through
model criticism~\citep{box1980sampling}, which we denote AVO-EB. While the
likelihood is intractable, it is fixed, thus one can view the optimization of
$\bfpsi$ through the lens of empirical Bayes, though the prior family
$p(\bftheta | \bfpsi)$ is not explicitly specified.

@@

\kcnote{Discuss hierarchical models and iid data x. Factor of $N/M$ for AVO-EB?}

@@

When $\rho$ is the Wasserstein distance, $\bftheta^*$ is referred to as the
minimum Wasserstein estimator. When the model is well specified, $\bftheta^*$
coincides with the true data-generating parameter; however, if the model is
misspecified, $\bftheta^*$ is typically different from the maximum likelihood
estimator (MLE), which is the minimizer of $KL(p_r(\bfx) | p(\bfx |
\bftheta)$~\citep{bernton2017inference}. In order to more effectively target
point estimates $\bftheta^*$, we augment Eqn.~\ref{eqn:vo-ug} with a
regularization term corresponding to the differential entropy $H$ of the
proposal distribution $q(\bftheta | \bfpsi)$. That is,
\begin{equation}
    U_g = \mathbb{E}_{\bftheta \sim q(\bftheta|\bfpsi)} [ {\cal L}_g ] + \gamma H(q(\bftheta|\bfpsi))
\end{equation}
where $\gamma \in \mathbb{R}^+$ is a hyper-parameter controlling the trade-off
between the generator objective and the tightness of the proposal distribution.
We refer to the algorithm with $\gamma>0$ as AVO-MWE.
For small values of $\gamma$,
proposal distributions with large entropy are not penalized, which results
in learning a smeared variation of the original simulator corresponding to the posterior predictive model. On the other hand,
for large values of $\gamma$, the procedure is encouraged to fit a proposal
distribution with low entropy, which has the effect of concentrating its density
tightly around one or a few $\bftheta$ values.
